---
title: "IMPC mouse and ortholog human gene counts"
author: "Jeremy Yang"
output:
  html_document:
    number_sections: yes
    toc: yes
---

```{r echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
base::date()
```

```{r}
library(readr)
library(data.table)
```

# IMPC from TCRD

<pre>
SELECT
        pt.protein_id,
        p.sym,
        p.uniprot,
        p.geneid,
        t.name,
        t.tdl,
        pt.id AS "tcrd_phenotype_id",
        pt.term_id AS "mp_term_id",
        pt.term_name AS "mp_term_name",
        pt.top_level_term_id AS "top_level_mp_term_id",
        pt.top_level_term_name AS "top_level_mp_term_name",
        pt.sex AS "mp_sex",
        pt.p_value,
        pt.percentage_change,
        pt.effect_size,
        pt.statistical_method
FROM
        phenotype pt
JOIN
        protein p ON p.id = pt.protein_id
JOIN
        t2tc ON t2tc.protein_id = pt.protein_id
JOIN
        target t ON t.id = t2tc.target_id
WHERE
        pt.ptype = 'IMPC'
</pre>

```{r message=FALSE}
tcrd_geno2pheno_impc <- read_delim("/home/data/TCRD/data/tcrd_geno2pheno_impc.tsv", "\t")
setDT(tcrd_geno2pheno_impc)
```
```{r echo=TRUE}
n_tcrd <- uniqueN(tcrd_geno2pheno_impc$protein_id)
sprintf("Human genes in TCRD IMPC g2p: %d", n_tcrd)
```

# IMPC from `metap` DB

<pre>
mouse.phen <- dbGetQuery(conn, "select * from mousephenotype")
mouse2human <- dbGetQuery(conn, "select h.protein_id human_protein_id,m.protein_id mouse_protein_id FROM homology h INNER JOIN homology m on h.homologene_group_id = m.homologene_group_id an
d h.tax_id = 9606 and m.tax_id = 10090")
mouse.phen <- merge(mouse.phen, mouse2human, by.x = "protein_id", by.y = "mouse_protein_id", sort = F)
write.table(mouse.phen, "mouse.phen.tsv", sep="\t", row.names=F)
</pre>

```{r}
mpml_mouse_phen <- read_delim("/home/data/metap/data/mouse.phen.tsv.gz", "\t")
setDT(mpml_mouse_phen)
```

```{r echo=TRUE}
sprintf("Mouse genes in MPML IMPC g2p: %d", uniqueN(mpml_mouse_phen$protein_id))
n_mpml <- uniqueN(mpml_mouse_phen$human_protein_id)
sprintf("Human genes in MPML IMPC g2p: %d", n_mpml)
```
```{r echo=TRUE}
sprintf("Human genes (TCRD) minus (MPML): (%d - %d) = %d (%.1f%%)", n_tcrd, n_mpml, n_tcrd-n_mpml, 100*(n_tcrd-n_mpml)/n_mpml)
```

# IMPC
## release 9.2 (Jan 29, 2019)
From ftp://ftp.ebi.ac.uk/pub/databases/impc/release-9.2/csv/ALL_genotype_phenotype.csv.gz
(Apparently includes {IMPC, MGP and EuroPhenome}_genotype_phenotype.csv.gz)

```{r message=FALSE}
impc_g2p <- read_delim("/home/data/IMPC/data/release-9.2/ALL_genotype_phenotype.csv", ",")
```

```{r echo=TRUE}
sprintf("Mouse genes in IMPC g2p: %d", uniqueN(impc_g2p$marker_accession_id))
```

# TCRD ortholog (mouse)

<pre>
SELECT
        id,
        protein_id,
        geneid,
        symbol AS "mgi_symbol",
        db_id AS "mgi_id",
        mod_url AS "mgi_url",
        taxid,
        species
FROM
        ortholog
WHERE
        taxid = 10090
</pre>

```{r}
ortholog <- read_delim("/home/data/TCRD/data/tcrd_ortholog_mouse.tsv", "\t", col_types="nccccccc")
setDT(ortholog)
ortholog[, c("id",  "taxid", "species", "mgi_url"):=NULL]
```

```{r echo=TRUE}
impc_g2p <- merge(impc_g2p, ortholog, by.x="marker_accession_id", by.y="mgi_id", all.x=T, all.y=F)
sprintf("Human genes in IMPC g2p, via orthology (protein_id): %d", uniqueN(impc_g2p$protein_id))
```
